---
title: "Analyse Macroinvertebrates"
author: "Jonathan Jupke"
date: "10 3 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(knitr)
library(kableExtra)
```
<!-- SETUP -->
```{r setup2}
source("R/setup.R")
source("R/fun_coef_plot.R")
```

Load all data.
```{r load data}
# saved TaxaAgg_NA_red from MACROINV script
data_abu = readRDS("data/taxa_bio.rds")
# saved EnvAgg_NA_red2 from MACROINV script 
data_env = readRDS("data/taxa_env.rds")
# both object just before rda in line 334 
# data with predictors pond and time 
data_exp = readRDS("data/taxa_exp.rds")
```

Scale and center environmental variables.
```{r prepare env data }
data_env %<>% scale(scale = TRUE, center =TRUE) %>% 
	as.data.frame()
```

Are there any co-linear variables left?
```{r}
corrplot::corrplot.mixed(corr = cor(data_env), lower = "shade", upper = "number")
```

Remove rare taxa.
```{r}
data_pa = data_abu %>% 
	mutate(row_id = rownames(data_abu)) %>% 
	pivot_longer(cols = !(row_id)) %>% 
	mutate(value2 = ifelse(value > 1, 1, value)) %>% 
	pivot_wider(id_cols = row_id,
	            values_from = value2,
	            names_from = name) %>% 
	dplyr::select(!row_id)

count_occ = colSums(data_pa)
rm_id = which(count_occ < 4)
data_abu = data_abu[,-rm_id]
```

Next we need to prepare the data. This includes transforming the response to a mvabund object and defining the contrasts.   
```{r}
# transform to mvabund format 
data_abu %<>% mvabund()
# define contrasts 
options(contrasts=c("contr.sum","contr.poly"))
data_exp$Season %<>% as.factor()
data_exp$Cat_Y %<>% as.factor()
contrasts(data_exp$Pond) = "contr.sum"
contrasts(data_exp$Season) = "contr.treatment"
contrasts(data_exp$Cat_Y) = "contr.treatment"
```

Create data sets for first and second year. 
```{r}
data_abu_y1 = data_abu[which(data_exp$Cat_Y == "1st"), ]
data_abu_y2 = data_abu[which(data_exp$Cat_Y == "2nd"), ]
data_env_y1 = data_env[which(data_exp$Cat_Y == "1st"), ]
data_env_y2 = data_env[which(data_exp$Cat_Y == "2nd"), ]
data_exp_y1 = filter(data_exp, Cat_Y == "1st")
data_exp_y2 = filter(data_exp, Cat_Y == "2nd")
```


# Fitting models 

Fit negative binomial models. 
```{r}
# For both years
mod_env = manyglm(data_abu ~ ., data = data_env)
mod_exp = manyglm(data_abu ~ Pond + Season + DaysFromBeg * Cat_Y, data = data_exp)
# For first year 
mod_env_y1 = manyglm(data_abu_y1 ~ ., data = data_env_y1)
mod_exp_y1 = manyglm(data_abu_y1 ~ Pond + Season + DaysFromBeg, data = data_exp_y1)
# For second year 
mod_env_y2 = manyglm(data_abu_y2 ~ ., data = data_env_y2)
mod_exp_y2 = manyglm(data_abu_y2 ~ Pond + Season + DaysFromBeg, data = data_exp_y2)
```

Check residuals of **env** models.
```{r}
walk(.x = 1:3, .f = ~ plot(mod_env, which = .x))
walk(.x = 1:3, .f = ~ plot(mod_env_y1, which = .x))
walk(.x = 1:3, .f = ~ plot(mod_env_y2, which = .x))
```

Check residuals of **exp** model.

```{r}
walk(.x = 1:3, .f = ~ plot(mod_exp, which = .x))
walk(.x = 1:3, .f = ~ plot(mod_exp_y1, which = .x))
walk(.x = 1:3, .f = ~ plot(mod_exp_y2, which = .x))
```

```{r eval = F}
mod_env_res = anova.manyglm(mod_env, 
	     p.uni = "adjusted", 
	     nBoot = 999)
mod_exp_res = anova.manyglm(mod_exp, 
	     p.uni = "adjusted", 
	     nBoot = 999)
mod_env_res_y1 = anova.manyglm(mod_env_y1,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_env_res_y2 = anova.manyglm(mod_env_y2,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_exp_res_y1 = anova.manyglm(mod_exp_y1,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_exp_res_y2 = anova.manyglm(mod_exp_y2,
	           p.uni = "adjusted",
	           nBoot = 999)
```


```{r eval = FALSE, echo = F}
mod_env_res = anova.manyglm(mod_env, 
	     p.uni = "adjusted", 
	     nBoot = 999)
mod_exp_res = anova.manyglm(mod_exp, 
	     p.uni = "adjusted", 
	     nBoot = 999)
mod_env_res_y1 = anova.manyglm(mod_env_y1,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_env_res_y2 = anova.manyglm(mod_env_y2,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_exp_res_y1 = anova.manyglm(mod_exp_y1,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_exp_res_y2 = anova.manyglm(mod_exp_y2,
	           p.uni = "adjusted",
	           nBoot = 999)
saveRDS(mod_env_res, "data/many_nb_env.rds")
saveRDS(mod_exp_res, "data/many_nb_exp.rds")
saveRDS(mod_env_res_y1, "data/many_nb_env_y1.rds")
saveRDS(mod_exp_res_y1, "data/many_nb_exp_y1.rds")
saveRDS(mod_env_res_y2, "data/many_nb_env_y2.rds")
saveRDS(mod_exp_res_y2, "data/many_nb_exp_y2.rds")
```

```{r echo = F}
mod_env_res = readRDS("data/many_nb_env.rds")
mod_exp_res = readRDS("data/many_nb_exp.rds")
mod_env_res_y1 = readRDS("data/many_nb_env_y1.rds")
mod_exp_res_y1 = readRDS("data/many_nb_exp_y1.rds")
mod_env_res_y2 = readRDS("data/many_nb_env_y2.rds")
mod_exp_res_y2 = readRDS("data/many_nb_exp_y2.rds")
```

## Results: Environmental covariates 

### both years 
```{r echo = F}
mod_env_res$table 	       	
```

### year 1
```{r echo = F}
mod_env_res_y1$table 	       	
```

### year 2
```{r echo = F}
mod_env_res_y2$table 	       	
```

## Results: Time and Pond 

### both years
```{r echo = F}
mod_exp_res$table 	       	
```
### year 1 
```{r echo = F }
mod_exp_res_y1$table 	       	
```
### year  2
```{r echo = F}
mod_exp_res_y2$table 	       	
```

## Coefficient plots: Environmental covariates 

### Both years 

```{r, echo = F}
plot_data = coef_plot(mod = mod_env)
colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

### Year 1 

```{r , echo = F}
plot_data = coef_plot(mod = mod_env_y1)
colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

### Year 2 

```{r, echo = F}
plot_data = coef_plot(mod = mod_env_y2)
colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

## Coefficient plots: Time and Pond 

### Both years 

```{r , echo = F}
plot_data = coef_plot(mod = mod_exp)

colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

### Year 1 

```{r, echo = F}
plot_data = coef_plot(mod = mod_exp_y1)

colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

### Year 2

```{r, echo = F}
plot_data = coef_plot(mod = mod_exp_y2)

colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

<!-- old stuff --> 

```{r eval = F}
walk(1:11,~ coefplot.manyglm(mod1,which.Xcoef = .x, cex.ylab= .3))
```
```{r eval = F}
#mod2 = gllvm(y = data4 , X = data5, formula = ~ ., family = "negative.binomial")
#saveRDS(mod2, "data/gllvm_nb.rds")
mod2 = readRDS("data/gllvm_nb.rds")
walk(.x = c(1,2,3), .f = ~ plot(mod2, which = .x))
```
```{r eval = F}
#mod3 = gllvm(y = data , X = data3, formula = ~ ., family = "ZIP")
#saveRDS(mod3, "data/gllvm_zip.rds")
mod3 = readRDS("data/gllvm_zip.rds")
walk(.x = c(1,2,3), .f = ~ plot(mod3, which = .x))
```

