---
title: "Analyse Macroinvertebrates"
author: "Jonathan Jupke"
date: "10 3 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(knitr)
library(kableExtra)
```
<!-- SETUP -->
```{r setup2}
source("R/setup.R")
source("R/fun_coef_plot.R")
```

Load all data.
```{r load data}
# saved TaxaAgg_NA_red from MACROINV script
data_abu = readRDS("data/taxa_bio.rds")
# saved EnvAgg_NA_red2 from MACROINV script 
data_env = readRDS("data/taxa_env.rds")
# both object just before rda in line 334 
# data with predictors pond and time 
data_exp = readRDS("data/taxa_exp.rds")
```

Scale and center environmental variables.
```{r prepare env data }
data_env %<>% scale(scale = TRUE, center =TRUE) %>% 
	as.data.frame()
```

Are there any co-linear variables left?
```{r}
corrplot::corrplot.mixed(corr = cor(data_env), lower = "shade", upper = "number")
```

Remove rare taxa.
```{r}
data_pa = data_abu %>% 
	mutate(row_id = rownames(data_abu)) %>% 
	pivot_longer(cols = !(row_id)) %>% 
	mutate(value2 = ifelse(value > 1, 1, value)) %>% 
	pivot_wider(id_cols = row_id,
	            values_from = value2,
	            names_from = name) %>% 
	dplyr::select(!row_id)

count_occ = colSums(data_pa)
rm_id = which(count_occ < 4)
data_abu = data_abu[,-rm_id]
```

Next we need to prepare the data. This includes transforming the response to a mvabund object and defining the contrasts.   
```{r}
# transform to mvabund format 
data_abu %<>% mvabund()
# define contrasts 
options(contrasts=c("contr.sum","contr.poly"))
data_exp$Season %<>% as.factor()
data_exp$Cat_Y %<>% as.factor()
contrasts(data_exp$Pond) = "contr.sum"
contrasts(data_exp$Season) = "contr.treatment"
contrasts(data_exp$Cat_Y) = "contr.treatment"
```

Create data sets for first and second year. 
```{r}
data_abu_y1 = data_abu[which(data_exp$Cat_Y == "1st"), ]
data_abu_y2 = data_abu[which(data_exp$Cat_Y == "2nd"), ]
data_env_y1 = data_env[which(data_exp$Cat_Y == "1st"), ]
data_env_y2 = data_env[which(data_exp$Cat_Y == "2nd"), ]
data_exp_y1 = filter(data_exp, Cat_Y == "1st")
data_exp_y2 = filter(data_exp, Cat_Y == "2nd")
```


# Fitting models 

Fit negative binomial models. 
```{r}
# For both years
mod_env = manyglm(data_abu ~ ., data = data_env)
mod_exp = manyglm(data_abu ~ Pond + Season + DaysFromBeg * Cat_Y, data = data_exp)
# For first year 
mod_env_y1 = manyglm(data_abu_y1 ~ ., data = data_env_y1)
mod_exp_y1 = manyglm(data_abu_y1 ~ Pond + Season + DaysFromBeg, data = data_exp_y1)
# For second year 
mod_env_y2 = manyglm(data_abu_y2 ~ ., data = data_env_y2)
mod_exp_y2 = manyglm(data_abu_y2 ~ Pond + Season + DaysFromBeg, data = data_exp_y2)
```

Check residuals of **env** models.
```{r}
walk(.x = 1:3, .f = ~ plot(mod_env, which = .x))
walk(.x = 1:3, .f = ~ plot(mod_env_y1, which = .x))
walk(.x = 1:3, .f = ~ plot(mod_env_y2, which = .x))
```

Check residuals of **exp** model.

```{r}
walk(.x = 1:3, .f = ~ plot(mod_exp, which = .x))
walk(.x = 1:3, .f = ~ plot(mod_exp_y1, which = .x))
walk(.x = 1:3, .f = ~ plot(mod_exp_y2, which = .x))
```

```{r eval = F}
mod_env_res = anova.manyglm(mod_env, 
	     p.uni = "adjusted", 
	     nBoot = 999)
mod_exp_res = anova.manyglm(mod_exp, 
	     p.uni = "adjusted", 
	     nBoot = 999)
mod_env_res_y1 = anova.manyglm(mod_env_y1,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_env_res_y2 = anova.manyglm(mod_env_y2,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_exp_res_y1 = anova.manyglm(mod_exp_y1,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_exp_res_y2 = anova.manyglm(mod_exp_y2,
	           p.uni = "adjusted",
	           nBoot = 999)
```


```{r eval = FALSE, echo = F}
mod_env_res = anova.manyglm(mod_env, 
	     p.uni = "adjusted", 
	     nBoot = 999)
mod_exp_res = anova.manyglm(mod_exp, 
	     p.uni = "adjusted", 
	     nBoot = 999)
mod_env_res_y1 = anova.manyglm(mod_env_y1,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_env_res_y2 = anova.manyglm(mod_env_y2,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_exp_res_y1 = anova.manyglm(mod_exp_y1,
	           p.uni = "adjusted",
	           nBoot = 999)
mod_exp_res_y2 = anova.manyglm(mod_exp_y2,
	           p.uni = "adjusted",
	           nBoot = 999)
saveRDS(mod_env_res, "data/many_nb_env.rds")
saveRDS(mod_exp_res, "data/many_nb_exp.rds")
saveRDS(mod_env_res_y1, "data/many_nb_env_y1.rds")
saveRDS(mod_exp_res_y1, "data/many_nb_exp_y1.rds")
saveRDS(mod_env_res_y2, "data/many_nb_env_y2.rds")
saveRDS(mod_exp_res_y2, "data/many_nb_exp_y2.rds")
```

```{r echo = F}
mod_env_res = readRDS("data/many_nb_env.rds")
mod_exp_res = readRDS("data/many_nb_exp.rds")
mod_env_res_y1 = readRDS("data/many_nb_env_y1.rds")
mod_exp_res_y1 = readRDS("data/many_nb_exp_y1.rds")
mod_env_res_y2 = readRDS("data/many_nb_env_y2.rds")
mod_exp_res_y2 = readRDS("data/many_nb_exp_y2.rds")
```

## Results: Environmental covariates 

### both years 
```{r echo = F}
mod_env_res$table 	       	
```

### year 1
```{r echo = F}
mod_env_res_y1$table 	       	
```

### year 2
```{r echo = F}
mod_env_res_y2$table 	       	
```

## Results: Time and Pond 

### both years
```{r echo = F}
mod_exp_res$table 	       	
```
### year 1 
```{r echo = F }
mod_exp_res_y1$table 	       	
```
### year  2
```{r echo = F}
mod_exp_res_y2$table 	       	
```

## Coefficient plots: Environmental covariates 

### Both years 

```{r, echo = F}
plot_data = coef_plot(mod = mod_env)
colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

### Year 1 

```{r , echo = F}
plot_data = coef_plot(mod = mod_env_y1)
colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

### Year 2 

```{r, echo = F}
plot_data = coef_plot(mod = mod_env_y2)
colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

## Coefficient plots: Time and Pond 

### Both years 

```{r , echo = F}
plot_data = coef_plot(mod = mod_exp)

colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

### Year 1 

```{r, echo = F}
plot_data = coef_plot(mod = mod_exp_y1)

colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

### Year 2

```{r, echo = F}
plot_data = coef_plot(mod = mod_exp_y2)

colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
plot.tas = levelplot(
  t(as.matrix(plot_data)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
print(plot.tas)
```

# Mixed model 
```{r}
library(lme4)
library(GLMMadaptive)
library(DHARMa)
```
```{r}
data_abu = as.data.frame(data_abu)
data_mixed = bind_cols(data_abu, data_env, data_exp)
data_mixed_y1 = filter(data_mixed, Cat_Y == "1st")
data_mixed_y2 = filter(data_mixed, Cat_Y == "2nd")
## -- are any species missing from year subsets? 

any(colSums(data_mixed[,1:ncol(data_abu)])== 0) 
any(colSums(data_mixed_y1[,1:ncol(data_abu)])== 0) 
any(colSums(data_mixed_y2[,1:ncol(data_abu)])== 0) 

## -- remove missing taxa 
data_mixed_y1 = data_mixed_y1[, - which(colSums(data_mixed_y1[,1:ncol(data_abu)])== 0)]
data_mixed_y2 = data_mixed_y2[, - which(colSums(data_mixed_y2[,1:ncol(data_abu)])== 0)]
```

```{r fitting mixed models, eval = F}
for (i in 1:ncol(data_abu)) {
	if (i == 1) models = list()
	models[[i]] = 
	try(
	glmer.nb(
		data_mixed[, i] ~ 
		Cond + O2  + NO3 
		+ PO4 + SO4 + NH4   
		+  (1 | Pond),
		data = data_mixed
	))
}
```

```{r fitting mixed models for year 1, eval = F}
## -- how many species are in data_mixed_y1? 
#names(data_mixed_y1)
## -- 16
for (i in 1:16) {
	if (i == 1) models_y1 = list()
	
	if (i != 14){
		models_y1[[i]] = 
	try(
	glmer.nb(
		data_mixed_y1[, i] ~ 
		Cond + O2  + NO3 
		+ PO4 + SO4 + NH4 
		+  (1 | Pond),
		data = data_mixed_y1
	))	
	}
	
}
```

```{r fitting mixed models for year 2, eval = F}
## -- how many species are in data_mixed_y2? 
names(data_mixed_y2)
## -- 
for (i in 1:23) {
	if (i == 1) models_y2 = list()


	models_y2[[i]] = 
	try(
	glmer.nb(
		data_mixed_y2[, i] ~ 
		Cond + O2  + NO3 
		+ PO4 + SO4 + NH4  + Season 
		+  (1 | Pond),
		data = data_mixed_y2
	))
}
```

Check variance inflation factor and residuals.
```{r vif mixed models both years, eval = F}
for (i in seq_along(models)){
	cond = class(models[[i]]) == "try-error"
	if (cond) next 
	print(paste(i, "-", round(max(vif(models[[i]])), 2)))
}
```

Check residuals for both years. 
```{r vif mixed models both years, eval = F}
simulateResiduals(fittedModel = models[[13]], plot = T)
# Problems with 1, 6, 12, 13, 14, 16, 17, 21, 22
```

Variance inflation factor for year 1.
```{r vif mixed models year 1, eval = F}
for (i in seq_along(models_y1)){
	cond = class(models_y1[[i]]) == "try-error"
	if (cond) next 
	print(paste(i, "-", round(max(vif(models_y1[[i]])), 2)))
}
## -- closer look at high VIF
vif(models_y1[[14]]) # Season, NH4, SO4, PO4, Cond are high  

```


```{r vif mixed models year 1, eval = F}
simulateResiduals(fittedModel = models_y1[[13]], plot = T)
```

```{r vif mixed models year 2, eval = F}
for (i in seq_along(models_y2)){
	cond = class(models_y2[[i]]) == "try-error"
	if (cond) next 
	print(paste(i, "-", round(max(vif(models_y2[[i]])), 2)))
}
simulateResiduals(fittedModel = models_y2[[13]], plot = T)
```


```{r plot mixed models, eval = T}
models = readRDS("../data/glmm_list.rds")
p_vec = c(1, 6, 12, 13, 14, 15,16, 17, 21, 22)
names(data_mixed)[p_vec]
re_ls = list()
fe_ls = list()
for (i in seq_along(models)){
	if (i %in% p_vec) next()
	lp_dt = ranef(models[[i]])
	lp_dt2 = fixef(models[[i]])
	ind = summary(models[[i]])
	
	re_ls[[i]] = data.table(pond = rownames(lp_dt$Pond),
		  sd   = lp_dt$Pond$`(Intercept)`,
		  taxon = names(data_mixed)[i])
	fe_ls[[i]] = data.table(variable = names(lp_dt2),
		  coefficient   = lp_dt2,
		  taxon = names(data_mixed)[i],
		  p_value = ind$coefficients[,4])

	if (i == length(models)) {
		re_ls %<>%
		rbindlist() %>%  
		pivot_wider(
			id_cols = taxon,
			names_from = pond,
			values_from = sd
		)
		
	}	
}
fe_ls %<>% rbindlist()
# holm's corrections 
n_test = nrow(fe_ls)
setorderv(fe_ls, "p_value")

fe_ls[, denominator := n_test:1]
fe_ls[, p_holm := p_value/denominator]
fe_ls[,stat_sig := p_holm <= 0.05]
fe_ls[stat_sig == TRUE,] %>%
		 
		pivot_wider(
			id_cols = taxon,
			names_from = variable,
			values_from = coefficient,
			values_fill = 0
		) -> 
	fe_ls2
colort   = colorRampPalette(c("blue", "white", "red"))
#a        = max(abs(plot_data))
r_lvlpdata = as.data.frame(re_ls)
f_lvlpdata = as.data.frame(fe_ls2)
rownames(r_lvlpdata) = r_lvlpdata$taxon
rownames(f_lvlpdata) = f_lvlpdata$taxon
r_lvlpdata = r_lvlpdata[,-1]
f_lvlpdata = f_lvlpdata[,-1]
r_plot.tas = levelplot(
  t(as.matrix(r_lvlpdata)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
f_plot.tas = levelplot(
  t(as.matrix(f_lvlpdata)),
  ylab = "",
  xlab  = "",
  col.regions = colort(100),
  at = seq(-2, 2, length = 100),
  scales = list(x = list(rot = 45))
)
#cowplot::plot_grid(r_plot.tas, f_plot.tas)
r_plot.tas
f_plot.tas
```
